---
title: "Assignment 6"
author: "Zhulin Yu"
date: "12/18/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load Libraries

```{r Load Libraries}
library(rpart)
```

Load Data

The MOOC data contains the following per-student variables:

certified (yes/no) - Whether or not a student paid for the course  

forum.posts (numeric) - How many forum posts a student made throughout the course

grade (numeric) - A student's average grade for the course exam

assignment (numeric) - A student's average grade for the course assignments

```{r}
#Upload the data sets MOOC1.csv and MOOC2.csv
M1 <- read.csv("MOOC1.csv")

M2 <- read.csv("MOOC2.csv")

```

# Decision tree
```{r}
#Using the rpart package generate a classification tree predicting certified from the other variables in the M1 data frame.

c.tree1 <- rpart(certified ~ forum.posts + grade + assignment, method="class", data=M1, control=rpart.control(minsplit=2, minbucket=1, cp=0.0001))

# Here, when using the defalt rpart setting, it cannot generate "leaves" and lead to "fit is not a tree, just a root" error. So I modified the fit setting to cp=0.002, which decrease the criterion to split data and also generate a number of branches.

#Check the results from the classifcation tree using the printcp() command

printcp(c.tree1)

#Plot your tree
plot(c.tree1)
post(c.tree1, file = "tree1.ps", title = "Certified: 1 - YES, 0 - NO")
```

The heading "xerror" in the printcp table stands for "cross validation error", it is the error rate of assigning students to certified/uncertified of the model averaged over 10-fold cross validation.

CP stands for "Cost Complexity" and represents the cost in error for adding a node to the tree. Notice it decreases as we add more nodes to the tree which implies that more nodes make better predictions.

However, more nodes also mean that we may be making the model less generalizable, this is known as "overfitting".

# Prune

If we are worried about overfitting we can remove nodes form our tree using the prune() command, setting cp to the CP value from the table that corresponds to the number of nodes we want the tree to terminate at. Let's set it to two nodes.

### However, maybe due to the nature of the raw data, two nodes cannot be achieved. I could only get 9 nodes by setting cp to 0.00392157. And this cp value reduces node number from 272 to 20.

```{r}
c.tree2 <- prune(c.tree1, cp = 0.00392157)
printcp(c.tree2)
#Visualize this tree and compare it to the one you generated earlier
plot(c.tree2)
post(c.tree2, file = "tree2.ps", title = "MOOC") #This creates a pdf image of the tree
```

### Now use both the original tree and the pruned tree to make predictions about the the students in the second data set. Which tree has a lower error rate?

```{r}
M2$predict1 <- predict(c.tree1, M2, type = "class")

M2$predict2 <- predict(c.tree2, M2, type = "class")

table(M2$certified, M2$predict1)

table(M2$certified, M2$predict2)

```

The pruned tree has a lower error rate: (239 + 2016)/10000 = 0.2255, compared to the original tree, whose error rate is: (1642 + 1702)/10000 = 0.3344.